---
title: "GARCH Modelling"
output: pdf_document
---

### Importing libraries

```{r, setup, include=FALSE}
library(rugarch)
library(ggplot2)
```

### Reading data

**Important:** Before running the code below, make sure your Knit directory is 'Document Directory'. This can be done by clicking the drop-down menu next to Knit, going to Knit directory and clicking on Document Directory.

```{r}
setwd("..")
#Inputting log-returns
sp_logret<-read.csv("Data/Processed/sp_logret.csv")$x
dow_logret<-read.csv("Data/Processed/dow_logret.csv")$x
nas_logret<-read.csv("Data/Processed/nas_logret.csv")$x

#Inputting residuals
sp_residuals<-read.csv("Data/Processed/sp_residuals.csv")$x
dow_residuals<-read.csv("Data/Processed/dow_residuals.csv")$x
nas_residuals<-read.csv("Data/Processed/nas_residuals.csv")$x

#Calculating AR
sp_ar <- sp_logret - sp_residuals
dow_ar <- dow_logret - dow_residuals
nas_ar <- nas_logret - nas_residuals
```

Separating data into train and test:

```{r}
sp_train_data_residuals <- sp_residuals[1:4000]
sp_test_data_residuals <- sp_residuals[4001:length(sp_residuals)]

dow_train_data_residuals <- dow_residuals[1:4000]
dow_test_data_residuals <- dow_residuals[4001:length(dow_residuals)]

nas_train_data_residuals <- nas_residuals[1:4000]
nas_test_data_residuals <- nas_residuals[4001:length(nas_residuals)]

sp_train_data_logret <- sp_logret[1:4000]
sp_test_data_logret <- sp_logret[4001:length(sp_logret)]

dow_train_data_logret <- dow_logret[1:4000]
dow_test_data_logret <- dow_logret[4001:length(dow_logret)]

nas_train_data_logret <- nas_logret[1:4000]
nas_test_data_logret <- nas_logret[4001:length(nas_logret)]
```

### Choosing the optimal parameters of GARCH(p,q):

We will choose the GARCH(p,q) model which has the lowest AIC. We will also be using the Student's t-distribution to model our volatility.

We will first define a function which we will run on our residuals data for each index. The function will go through a series of GARCH models and output the one with the lowest AIC.

```{r}
opm_garch <- function(x){
  final.aic <- Inf
  final.order <- c(0,0)
  for (i in 1:4) for (j in 1:4){
    x_model <-  ugarchspec(variance.model=list(garchOrder=c(i,j)),
                mean.model=list(armaOrder=c(0,0)), distribution.model = "std")
    current.aic <- infocriteria(ugarchfit(spec=x_model,data=x,solver='hybrid'))[1]
    if (current.aic < final.aic){
      final.aic <- current.aic
      final.order <- c(i,j)
  }
}
final.order
}
```

The below code is commented out as the algorithm takes some time to run and we don't need to run it as we already have obtained our previous parameters.

```{r}
#opm_garch(sp_train_data)
#opm_garch(dow_train_data)
#opm_garch(nas_train_data)
```

The optimizer chose GARCH(2,1) to model the volatility for all of our indexes.

Hence, we have the following models for each index's log-returns:

-   **sp:** AR(1)+GARCH(1,2)

-   **dow:** AR(1)+GARCH(1,2)

-   **nas:** ARMA(3,2)+GARCH(1,2)

We will now define our models for each index:

```{r}
sp.model = ugarchspec(variance.model = list(model = 'sGARCH' , garchOrder = c(2,1)),mean.model = list(armaOrder = c(0 , 0)))

dow.model = ugarchspec(variance.model = list(model = 'sGARCH' , garchOrder = c(2,1)),mean.model = list(armaOrder = c(0 , 0)))

nas.model = ugarchspec(variance.model = list(model = 'sGARCH' , garchOrder = c(2,1)),mean.model = list(armaOrder = c(0 , 0)))
```

### Fitting our models:

Now we will fit our model on training data and output the coefficients:

```{r}
sp.model.fit <- ugarchfit(sp.model, sp_train_data, solver = 'solnp')

dow.model.fit <- ugarchfit(dow.model, dow_train_data, solver = 'solnp')

nas.model.fit <- ugarchfit(nas.model, nas_train_data, solver = 'solnp')

options(scipen = 999)
sp.model.fit@fit$matcoef
dow.model.fit@fit$matcoef
nas.model.fit@fit$matcoef
```

Our GARCH(p,q) models take the following shape:

[**S&P500:**]{.ul}

$$
\sigma^{2}_t =  0.0000030 + 0.848\sigma^{2}_{t-1} + 0.047\epsilon^{2}_{t-1} + 0.079\epsilon^{2}_{t-2} 
$$

[**Dow Jones:**]{.ul}

$$
\sigma^{2}_t =  0.0000022 + 0.859\sigma^{2}_{t-1} + 0.039\epsilon^{2}_{t-1} + 0.085\epsilon^{2}_{t-2} 
$$

[**NASDAQ:**]{.ul}

$$
\sigma^{2}_t =  0.0000027 + 0.882\sigma^{2}_{t-1} + 0.022\epsilon^{2}_{t-1} + 0.084\epsilon^{2}_{t-2} 
$$

Let us also produce some plots to compare our fitted values against the squared residuals:

```{r}
ggplot() + 
  geom_line(aes(x=(1:length(sp_train_data)), y = (sp_train_data)^2, color = "Squared Residuals")) +
  geom_line(aes(x=(1:length(sp_train_data)),y=sp.model.fit@fit[["sigma"]], color = "Fitted Values"))+
  labs(x = "Index", y = "", color = "Legend", title = 'Plot of squared residuals against fitted for S&P') +
  scale_colour_manual(values = c("Squared Residuals"="black", "Fitted Values"="red"))
                                 
ggplot() + 
  geom_line(aes(x=(1:length(dow_train_data)), y = (dow_train_data)^2, color = "Squared Residuals")) +
  geom_line(aes(x=(1:length(dow_train_data)),y=dow.model.fit@fit[["sigma"]], color = "Fitted Values"))+
  labs(x = "Index", y = "", color = "Legend", title = 'Plot of squared residuals against fitted for Dow') +
  scale_colour_manual(values = c("Squared Residuals"="black", "Fitted Values"="red"))

ggplot() + 
  geom_line(aes(x=(1:length(nas_train_data)), y = (nas_train_data)^2, color = "Squared Residuals")) +
  geom_line(aes(x=(1:length(nas_train_data)),y=nas.model.fit@fit[["sigma"]], color = "Fitted Values"))+
  labs(x = "Index", y = "", color = "Legend", title = 'Plot of squared residuals against fitted for NASDAQ') +
  scale_colour_manual(values = c("Squared Residuals"="black", "Fitted Values"="red"))
```

### Forecasting on our model:

Now we will forecast on our data. The model forecasts one-day ahead and we will be refitting our model every 100 days:

```{r}
sp.model.forecast <- ugarchroll(spec = sp.model , data = sp_residuals, n.start = 4000, refit.every = 100, refit.window = 'moving', solver='solnp')

dow.model.forecast <- ugarchroll(spec = dow.model , data = dow_residuals, n.start = 4000, refit.every = 100, refit.window = 'moving', solver='solnp')

nas.model.forecast <- ugarchroll(spec = dow.model , data = nas_residuals , n.start = 4000, refit.every = 100, refit.window = 'moving', solver='solnp')
```

Let us produce some plots to check how our model does

```{r}
ggplot() + 
  geom_line(aes(x=(1:length(sp_test_data)), y = (sp_test_data)^2, color = "Squared Residuals")) +
  geom_line(aes(x=(1:length(sp_test_data)),y=sp.model.forecast@forecast$density[,'Sigma'], color = "Fitted Values"))+
  labs(x = "Index", y = "", color = "Legend", title = 'Plot of squared residuals against forecast for S&P') +
  scale_colour_manual(values = c("Squared Residuals"="black", "Fitted Values"="red"))
                                 
ggplot() + 
  geom_line(aes(x=(1:length(dow_test_data)), y = (dow_test_data)^2, color = "Squared Residuals")) +
  geom_line(aes(x=(1:length(dow_test_data)),y=dow.model.forecast@forecast$density[,'Sigma'], color = "Fitted Values"))+
  labs(x = "Index", y = "", color = "Legend", title = 'Plot of squared residuals against forecast for Dow') +
  scale_colour_manual(values = c("Squared Residuals"="black", "Fitted Values"="red"))

ggplot() + 
  geom_line(aes(x=(1:length(nas_test_data)), y = (nas_test_data)^2, color = "Squared Residuals")) +
  geom_line(aes(x=(1:length(nas_test_data)),y=nas.model.forecast@forecast$density[,'Sigma'], color = "Fitted Values"))+
  labs(x = "Index", y = "", color = "Legend", title = 'Plot of squared residuals against forecast for NASDAQ') +
  scale_colour_manual(values = c("Squared Residuals"="black", "Fitted Values"="red"))
```

### Variance at Risk (Var):

In order to compare our LSTM with GARCH, we will calculate the VaR and see how well both models do.

Let's start of a with a quick definition of VaR:

*'Value at risk (VaR) is a measure of the risk of loss for investments. It estimates how much a set of investments might lose (with a given probability), given normal market conditions, in a set time period such as a day. VaR is typically used by firms and regulators in the financial industry to gauge the amount of assets needed to cover possible losses.' ([Wikipedia](https://en.wikipedia.org/wiki/Value_at_risk))*

As seen before, we will model our returns according to the Student's t-distribution. Hence we will calculate VaR daily using the following formula:

$$
VaR_{t|t-1}(\alpha) = \mu_{t} + \hat{\sigma}_{t|t-1} * F^{-1}(\alpha)
$$

where $\hat{\sigma}_{t|t-1}$ is the conditional standard deviation given the information at $t-1$, $F^{−1}(\alpha)$ is the inverse PDF function of t-distribution at the significance level $\alpha$ and $\mu_t$ is the value of the mean equation at time $t$

In order to test our VaR estimates we will use a measure introduced in Kupiec (1995).

Let $N=\sum_{t=1}^{T}I_t$ where $I_t$ is $1$ if the actual log-return at time $t+1$ is greater than $VaR_{t+1|t}(\alpha)$ and $0$ otherwise over $T$ days. It was argued Kupiec (1995) that $N\sim Bin(T,alpha)$ i.e. N is binomially distributed. Hence we would expect $N$ to be around $T*\alpha$.

Let us first define a function which we will use to calculate VaR:

```{r}
VaR <- function(x.model,x_train_data_residuals,x_test_data_residuals,x_test_data_logret,x_ar,alpha){
  x_shape <- fitdist(distribution = 'std' , x = x_train_data_residuals)$pars[3]
  n = ceiling(length(x_test_data_residuals)/100)
  x_list <- rep(0,times=n)
  for (i in (1:n)){
    mean = na.omit(x_ar[4001:length(x_ar)][(100*(i-1)+1):(100*i)])
    sigma = na.omit(x.model@forecast$density[,'Sigma'][(100*(i-1)+1):(100*i)])
    VaR95_td <- mean+sigma*qdist(distribution='std', shape=x_shape, p=alpha)
    x_list[i] <- sum(na.omit(x_test_data_logret[(100*(i-1)+1):(100*i)]) < VaR95_td)
  }
  return(x_list)
}
```

We will break our test data into periods of 100 days and our significance level at 5%. Hence $T=100$ and $\alpha = 0.05$. We will then take the average of all the periods. We expect the average to be around $5$. The 95% confidence interval is $[1,10]$ so we would expect $N$ to lie within this range.

```{r}
sp_list <- VaR(sp.model.forecast,sp_train_data_residuals,sp_test_data_residuals,sp_test_data_logret,sp_ar,0.05)
dow_list <- VaR(dow.model.forecast,dow_train_data_residuals,dow_test_data_residuals,dow_test_data_logret,dow_ar,0.05)
nas_list <- VaR(nas.model.forecast,nas_train_data_residuals,nas_test_data_residuals,nas_test_data_logret,nas_ar,0.05)

mean(sp_list[1:10])
mean(dow_list[1:12])
mean(nas_list[1:12])
```

As we see, all our indexes lie within this range hence our GARCH models seems to be a good fit in modelling our VaR.

### Expected Shortfall (ES):

We will now calculate the ES estimates of our model. Let us begin with a short introduction:

*Expected shortfall (ES) is a risk measure---a concept used in the field of financial risk measurement to evaluate the market risk or credit risk of a portfolio. The "expected shortfall at q% level" is the expected return on the portfolio in the worst q% of cases. ES is an alternative to value at risk that is more sensitive to the shape of the tail of the loss distribution. ([Wikipedia](https://en.wikipedia.org/wiki/Expected_shortfall))*

Practically, you can think of it as taking the average of the q% of cases.

Given that we know the VaR, we can calculate ES using the following equation:

$$
\frac{1}{\alpha} \int_{0}^{\alpha} VaR(\alpha) \,d\alpha
$$

We will use a similar method as VaR i.e. we split our test data into 100 days, calculate ES for every day and then average the 100 days.

```{r}
VaRalpha <- function(x.model,x_train_data,x_test_data,x_ar,alpha,j,i){
        x_shape <- fitdist(distribution = 'std' , x = x_train_data)$pars[3]
        mean = x_ar[4001:length(x_ar)][(100*(i-1)+1):(100*i)]
        sigma = x.model@forecast$density[,'Sigma'][(100*(i-1)+1):(100*i)]
        VaR95_td <- mean[j]+sigma[j]*qdist(distribution='std', shape=x_shape, p=alpha)
}


ES <- function(x.model,x_train_data,x_test_data,x_ar,alpha){
  n = ceiling(length(x_test_data)/100)
  x_list <- rep(0,times=n)
  for (i in (1:n)){
    if (i==n){
      x_integrate <- rep(0,length(x_test_data)-(100)*(i-1))
      for (j in 1:(length(x_integrate))){
        x_integrate[j] <- as.numeric(integrate(VaRalpha,lower=0,upper=alpha,x.model=x.model,x_train_data=x_train_data,x_test_data=x_test_data,x_ar=x_ar,j=j,i=i)[1])
        x_list[i] <- mean(x_integrate)
      }
    } 
    else {
      x_integrate <- rep(0,(100))
      for (j in 1:(length(x_integrate))){
        x_integrate[j] <- as.numeric(integrate(VaRalpha,lower=0,upper=alpha,x.model=x.model,x_train_data=x_train_data,x_test_data=x_test_data,x_ar=x_ar, j=j, i=i)[1])/(alpha)
        x_list[i] <- mean(x_integrate)
      }
    }
    print(i)
}
return(x_list)
}
```

```{r}
setwd('..')

#sp_list_ES <- ES(sp.model.forecast,sp_train_data,sp_test_data,sp_ar,0.05)
#write.csv(sp_list_ES, 'Data/Processed/sp_list_ES.csv', row.names=T)

#dow_list_ES <- ES(dow.model.forecast,dow_train_data,dow_test_data,dow_ar,0.05)
#write.csv(dow_list_ES, 'Data/Processed/dow_list_ES.csv', row.names=T)

#nas_list_ES <- ES(nas.model.forecast,nas_train_data,nas_test_data,dow_ar,0.05)
#write.csv(nas_list_ES, 'Data/Processed/nas_list_ES.csv', row.names=T)

sp_residuals<-read.csv("Data/Processed/sp_list_ES.csv")$x
dow_residuals<-read.csv("Data/Processed/dow_list_ES.csv")$x
nas_residuals<-read.csv("Data/Processed/nas_list_ES.csv")$x
```

We will compute the actual ES of the test data by averaging all the log-return prices which are less than 5% quantile.

```{r}
ES_actual <- function(x_test_data,alpha){
  n <- ceiling(length(x_test_data)/100)
  x_list <- rep(0,times=n)
  for (i in (1:n)){
    if (i==n){
      x_data <- x_test_data[(100*(i-1)+1):length(x_test_data)]
      x_list[i] <- mean(x_data[x_data <= quantile(x_data,alpha)])
    } 
    else {
      x_data <- x_test_data[(100*(i-1)+1):(100*i)]
      x_list[i] <- mean(x_data[x_data <= quantile(x_data,alpha)])
    }
  }
  return(x_list)
}
```

```{r}
sp_list_ES_actual <- ES_actual(sp_test_data,0.05)
dow_list_ES_actual <- ES_actual(dow_test_data,0.05)
nas_list_ES_actual <- ES_actual(nas_test_data,0.05)
```

We calculate the difference between the ES our model predicted and actual ES.

```{r}
sp_ES_change <- mean((sp_list_ES_actual - sp_list_ES)^2)
dow_ES_change <- mean((dow_list_ES_actual - dow_list_ES)^2)
nas_ES_change <- mean((nas_list_ES_actual - nas_list_ES)^2)
```
